Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-16 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:16	08:26:02	(Fri16.Oct)

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. In sonar systems with less than 32 channels the complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. For larger systems the inversion step dominates the computation, the use of approximations through space reduction techniques become attractive. Some examples of such techniques include beamspace proessing on a GPU~\cite{Asen2013}, the closely related principal component analysis (PCA) method~\cite{Kim2014}, or to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. At best the mentioned methods are able to reduce the complexity to O($M^2$).

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g \emph{et al}. in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams\todo{ha med at i US \o{}nsker man ulike vindus-typer p\aa{} forskjellig range?}. LCA applies a set of predefined windows, or tapers, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a an multi-taper technique. One such method is described by Stankwitz \emph{et al}. in radar imaging~\cite{Stankwitz1995}. Here the best of 2 or 3 windows are selected, but unlike with LCA no micro-steered variations are used.

Synnev\aa{}g \emph{et al}. suggested using a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were micro-steered variants of a fairly wide Kaiser window. While achieving good results, it was not made clear how LCA performs with more or fewer windows, and how they should be steered. Also, it is not clear how LCA performs in systems with unfocused transmit beams.

>>>>

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. In sonar systems with less than 32 channels the complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. For larger systems the inversion step dominates the computation. This can be dealt with by approximating the full covariance matrix with a small one using space reduction techniques such as beamspace processing~\cite{Asen2013}, or the closely related principal component analysis (PCA) method~\cite{Kim2014}. Another method is to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. However, at best the mentioned methods are able to reduce the complexity to O($M^2$).

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g \emph{et al}. in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams. LCA applies a set of predefined windows, or tapers, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a multi-taper technique. One such method is described by Stankwitz \emph{et al}. in radar imaging~\cite{Stankwitz1995}, where the best out of 2 or 3 windows are selected. However, LCA is different in that it typically selects from more windows and allow micro-steered variations of them. This gives LCA an adaptive freedom that is somewhere in between that of traditional multi-taper techniques and MVDR.

Synnev\aa{}g \emph{et al}. suggested using a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were micro-steered variants of a fairly wide Kaiser window, and with the rectangular window included. This was motivated by the desire to use wide windows with maximum sidelobe suppression near the receiver, and narrow windows at greater depths for maximum sensitivity and depth penetration. While achieving good results, it was not made clear how how many windows LCA needs, or how they ideally should be steered. Also, it is not apparent how LCA performs in sonar systems where the transmit beam is unfocused.
